Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

	\item Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.  For modest $d$, the possible sample sizes, $n$, are quite sparse, as is shown in Figure \ref{fig:keister-err}. Choosing the nodes to lie on a grid corresponds to a tensor product midpoint cubature rule, which would normally be expected to have an error of $\mathcal{O}(n^{-2/d})$  for general $d$.  The error decay of $\mathcal{O}(n^{-1/5})$ rather than $\mathcal{O}(n^{-2/6})$ for this example may be due to the lack of smoothness of $f$ near the boundaries of the unit cube.
	\item Simple MC (orange) are a substantial improvement over grid nodes.  Their root mean squared error is known to decay as $\mathcal{O}(n^{-1/2})$, as is observed here.  As a reminder, for IID nodes the root mean squared error
	\begin{equation}\label{eq:IIDerror}
		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
	\end{equation*}
	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  For simple MC
\end{itemize}

>>>>

	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.  For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.

	Choosing the nodes to lie on a grid corresponds to a tensor product midpoint cubature rule, which would normally be expected to have an error of $\mathcal{O}(n^{-2/d})$  for general $d$.  The error decay of $\mathcal{O}(n^{-1/5})$ rather than $\mathcal{O}(n^{-2/6})$ for this example may be due to the lack of smoothness of $f$ near the boundaries of the unit cube.

	\item \emph{IID nodes.}  Simple MC (orange) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
	\begin{equation}\label{eq:IIDerror}
		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
	\end{equation}
	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}. For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.

	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude for large enough $n$.  Typically, low discrepancy sequences have preferred sample sizes.  For this Sobol' sequence, $n$ an positive integer power of $2$ tends to give better accuracy than arbitrary $n$.

\end{itemize}
